Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2020-11-04 summary.Rmd
added algorithms for ssc and mm

Both Algorithm 2 and Sparse Subspace Clustering produce an affinity matrix $B$ 
for which $B_{ij} \to 0$ if vertices $i$ and $j$ belong in different 
communities. Spectral clustering was performed on $B$ to obtain the final 
community estimates.

For estimating $\{\lambda^{(kl)}\}_K$, both methods usethe ground truth community 
labels.

>>>>

Both Algorithm 2 and Sparse Subspace Clustering produce an affinity matrix $B$ 
for which $B_{ij} \to 0$ if vertices $i$ and $j$ belong in different 
communities. Spectral clustering was performed on $B$ to obtain the final 
community estimates. More specifically, clustering was performed on a 
normalized Laplacian eigenmap of dimension $K + 1$ by fitting Gaussian Mixture 
Model.

For estimating $\{\lambda^{(kl)}\}_K$, both methods use the ground truth 
community labels.

### Sparse subspace Clustering and Modularity Maximization

**TODO** find citations for these

For each simulation, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
from a joint distribution, the edge probability matrix $P$ is constructed using 
the popularity vectors, and finally an unweighted and undirected adjacency 
matrix $A$ is drawn from $P$. Three community detection methods are applied to 
$A$: Algorithm 2 from this paper, Sparse Subspace Clustering, and Modularity 
Maximization. The latter two algorithms are detailed as follows:

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering.}
  \KwData{Adjacency matrix $A$, number of communities $K$.}
  \KwResult{Community assignments $1, ..., K$.}
  Embed $A$ in $d$ dimensions using ASE to obtain 
  $\hat{Y} \in \mathbb{R}^{n \times d}$.\;
  Fit a sparse linear regression model to each row of $\hat{Y}$ using the 
  remaining $n - 1$ rows of $Y$ as predictors. The models must exclude the 
  intercept term.\;
  Construct $n \times n$ symmetric matrix $C$ such that $c_{ij}$ is the $j^{th}$
  coefficient for the $i^{th}$ model.\;
  Normalize the columns of $|C|$ to obtain $\tilde{C}$.\;
  Compute the affinity matrix $B = \tilde{C} + \tilde{C}^\top$.\;
  Construct graph $G$ using $B$ as its similarity matrix.\;
  Partition $G$ into $K$ disconnected components (e.g., using edge thresholding 
  or spectral clustering).\;
  Map each partition to the community labels $1, ..., K$\;
\end{algorithm}

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Modularity Maximization (Louvain Method)}
  \KwData{Adjacency matrix $A$, number of communities $K$.}
  \KwResult{Community assignments $1, ..., K$}
  Initialize the clustering by assigning to each vertex a unique community 
  label.\;
  For each vertex $v_i$, reassign the label to match one of its neighboring 
  vertices, and compute the change in modularity for each reassignment.\;
  Reassign $v_i$ to the community that results in the most positive change in 
  modularity, if possible. If modularity does not increase by changing the 
  label, skip the vertex.\;
  Group vertices with the same label and treat as one vertex.\;
  Repeat steps (3) and (4) until termination or the desired number of 
  communities is estimated.
\end{algorithm}
